---
title: "Code for Figures in Manuscript"
author: "F. King, A.L. Prendergast, E. Frost, G.L. Stannard, J. Garvey, R. Cosgrove, R. Drysdale, N. Hausmann"
output: 
  pdf_document:
    latex_engine: xelatex
---

# Libraries 
```{r Libraries,echo=FALSE,message = FALSE,warning = FALSE}
knitr::opts_chunk$set(echo=FALSE,message = FALSE,warning = FALSE)

{
  pacman::p_load(
    here,
    janitor,
    tidyverse,lubridate,dtw,
    cowplot,ggx,viridis,ShellChron,
    patchwork,
    RColorBrewer) 
  
  theme_set(theme_cowplot())
}

# This function extracts statistical attributes
lmp <- function (modelobject) {
    if (class(modelobject) != "lm") stop("Not an object of class 'lm' ")
    f <- summary(modelobject)$fstatistic
    p <- pf(f[1],f[2],f[3],lower.tail=F)
    attributes(p) <- NULL
    return(p)
}


```

# Data
```{r}

# Isotope data
Isotopes <- read.csv(here("data","Isotope_data.csv"))

# LIBS Data
LIBS <- read.csv(here("data","LIBS_data.csv"))

# SST Data
SST <- read.csv(here("data","SST_Data-Pinmatik.csv")) %>% 
  mutate(season=as_factor(season) %>% fct_shift(3),
         date=as.Date(date))

# Chlorophyll Data
Chl <- read_csv(here("data","chl.csv"), col_types = cols(time = col_character()),skip = 6) %>% 
  mutate(time=lubridate::as_date(time),
         time=round_date(time,unit="month"),
         month=as_factor(month(time)),
         year= as_factor(year(time)))


```

# Figure 1 - Time series of Chlorophyll-a data
```{r}

Chl %>%
  group_by(month) %>%
  summarize(avg_chl = mean(chl, na.rm = TRUE),
            sd_chl = sd(chl, na.rm = TRUE)) %>% 
ggplot()+
  aes(x = month, y = avg_chl, group=1) +
  geom_path(col="darkgreen",linewidth=1) +
  geom_ribbon(aes(ymin = avg_chl - sd_chl, ymax = avg_chl + sd_chl),
              alpha = 0.3,fill="darkgreen") +
  xlab("Month") +
  ylab(expression("Average mass concentration of chlorophyll-a (mg/m"^3*")")) +
  ylim(0,3)+
  ggtitle("Monthly Changes in Chlorophyll-a", subtitle= "Period of 1997 - 2023")+
  theme_cowplot(20)

# ggsave(here("Figures", "Figure 1.png"), units = "cm", height = 29.7, width = 20.7)

```


# Figure 5 - Elemental maps and line-scans of *C. solida* specimens from Rocky Cape South
## Individual graphs
### C07
```{r}

C07_Map <- LIBS %>% 
  filter(specimen=="C07",
         sample_mode=="map",
         mg_ca<0.6) %>% 
  ggplot() +
  aes(x/ 1000,
      y/ 1000,
      col = mg_ca) +
  geom_point(size = 1,show.legend = FALSE) +
  scale_color_viridis(option = "G",
                      begin = 0,
                      end = 0.95) +
  geom_path(data = LIBS %>% filter(specimen=="C07",sample_mode=="trace"),
            aes(x = x/ 1000, y = y/ 1000),
            col = "red",
            linetype="dashed",
            size=0.7)+
  coord_fixed()+  
  scale_x_continuous(breaks=c(15,14,13),trans = "reverse")+
    ylab("y-axis [mm]")+ggtitle("C7")+
  xlab("x-axis [mm]")+labs(col = "Mg/Ca [a.u.]")



C07_Line <- LIBS %>% 
  filter(specimen=="C07",
         sample_mode == "trace") %>%
  ggplot() +
  aes(dist,
      mg_ca,
      col = mg_ca) +
  geom_path(show.legend=FALSE,size=1) +
  scale_color_viridis(option = "G",
                      begin = 0.5,
                      end = 0.95) +
  ylab("Mg/Ca [a.u.]") +
  xlab("Distance to edge [mm]")


```

### C08
```{r C08}

C08_Map <- LIBS %>% 
  filter(specimen=="C08",
         sample_mode=="map",
         mg_ca<0.6,
         x>29000) %>% 
  ggplot() +
  aes(x/ 1000,# transofrm to mm-scale
      y/ 1000,# transofrm to mm-scale
      col = mg_ca) + 
  geom_point(size = 1.5,show.legend = FALSE) +
  scale_color_viridis(option = "G",
                      begin = 0,
                      end = 0.95) +
  coord_fixed()+  scale_x_reverse()+

  geom_path(data = LIBS %>% filter(specimen=="C08",sample_mode=="trace"),
            aes(x = x/ 1000, y = y/ 1000),
            col = "red",
            linetype="dashed",
            size=0.7)+
    ylab("y-axis [mm]")+ggtitle("C8")+
  xlab("x-axis [mm]")+labs(col = "Mg/Ca [a.u.]")



C08_Line <- LIBS %>% 
  filter(specimen=="C08",
         sample_mode == "trace") %>%
  ggplot() +
  aes(dist,
      mg_ca,
      col = mg_ca) +
  geom_path(show.legend=FALSE,size=1) +
  scale_color_viridis(option = "G",
                      begin = 0.5,
                      end = 0.95) +
  ylab("Mg/Ca [a.u.]") +
  xlab("Distance to edge [mm]")


```

### C09
```{r C09}

C09_Map <- LIBS %>% 
  filter(specimen=="C09",
         sample_mode=="map",
         mg_ca<0.6,
         x>21000) %>% 
  ggplot() +
  aes(x/ 1000,
      y/ 1000,
      col = mg_ca) +
  geom_point(size = 1) +
  scale_color_viridis(option = "G",
                      begin = 0,
                      end = 0.95) +
  coord_fixed()+  scale_x_reverse()+

  geom_path(data = LIBS %>% filter(specimen=="C09",sample_mode=="trace"),
            aes(x = x/ 1000, y = y/ 1000),
            col = "red",
            linetype="dashed",
            size=0.7)+
    ylab("y-axis [mm]")+ggtitle("C9")+
  xlab("x-axis [mm]")+labs(col = "Mg/Ca [a.u.]")



C09_Line <- LIBS %>% 
  filter(specimen=="C09",
         sample_mode == "trace") %>%
  ggplot() +
  aes(dist,
      mg_ca,
      col = mg_ca) +
  geom_path(show.legend=FALSE,size=1) +
  scale_color_viridis(option = "G",
                      begin = 0.5,
                      end = 0.95) +
  ylab("Mg/Ca [a.u.]") +
  xlab("Distance to edge [mm]")



```

## Final Plot Fig 5
```{r}

C07_Map+C08_Map+C09_Map+C07_Line+C08_Line+C09_Line+plot_annotation(
                          title="South",
                          tag_levels = "a"
                          )+plot_layout(ncol = 3,widths = c(0.5,0.5),guides = "collect")

# ggsave(here("Figures", "Figure 5.png"), units = "cm", width = 29.7, height = 14.85)

```





# Figure 6 - Elemental maps and line-scans of *C. solida* specimens from Rocky Cape North
## Individual Graphs
### C16
```{r }

C16_Map <- LIBS %>% 
  filter(specimen=="C16",
         sample_mode=="map",
         mg_ca<0.6,
         x>29000) %>% 
  ggplot() +
  aes(x/ 1000,
      y/ 1000,
      col = mg_ca) +
  geom_point(size = 1) +
  scale_color_viridis(option = "G",
                      begin = 0,
                      end = 0.95) +
  coord_fixed()+  scale_x_reverse()+

  geom_path(data = LIBS %>% filter(specimen=="C16",sample_mode=="trace"),
            aes(x = x/ 1000, y = y/ 1000),
            col = "red",
            linetype="dashed",
            size=0.7)+
      ylab("y-axis [mm]")+ ggtitle("C16")+
  xlab("x-axis [mm]")+labs(col = "Mg/Ca [a.u.]")



C16_Line <- LIBS %>% 
  filter(specimen=="C16",
         sample_mode == "trace") %>%
  ggplot() +
  aes(dist,
      mg_ca,
      col = mg_ca) +
  geom_path(show.legend=FALSE,size=1) +
  scale_color_viridis(option = "G",
                      begin = 0.5,
                      end = 0.95) +
  ylab("Mg/Ca [a.u.]") +
  xlab("Distance to edge [mm]")

  


```

### C17
```{r}

C17_Map <- LIBS %>% 
  filter(specimen=="C17",
         sample_mode=="map",
         mg_ca<0.6,
         x>25000) %>% 
  ggplot() +
  aes(x/ 1000,
      y/ 1000,
      col = mg_ca) +
  geom_point(size = 1,show.legend=FALSE) +
  scale_color_viridis(option = "G",
                      begin = 0,
                      end = 0.95) +
  geom_path(data = LIBS %>% filter(specimen=="C17",sample_mode=="trace"),
            aes(x = x/ 1000, y = y/ 1000),
            col = "red",
            linetype="dashed",
            size=0.7)+
  coord_fixed()+
  scale_x_reverse()+
    ylab("y-axis [mm]")+ggtitle("C17")+
  xlab("x-axis [mm]")+labs(col = "Mg/Ca [a.u.]")



C17_Line <- LIBS %>% 
  filter(specimen=="C17",
         sample_mode == "trace") %>%
  ggplot() +
  aes(dist,
      mg_ca,
      col = mg_ca) +
  geom_path(show.legend=FALSE,size=1) +
  scale_color_viridis(option = "G",
                      begin = 0.5,
                      end = 0.95) +
  ylab("Mg/Ca [a.u.]") +
  xlab("Distance to edge [mm]")

  



```

### C18
```{r}

C18_Map <- LIBS %>% 
  filter(specimen=="C18",
         sample_mode=="map",
         mg_ca<0.6,
         x>25000) %>% 
  ggplot() +
  aes(x/ 1000,
      y/ 1000,
      col = mg_ca) +
  geom_point(size = 1,show.legend=FALSE) +
  scale_color_viridis(option = "G",
                      begin = 0,
                      end = 0.95) +
  coord_fixed()+  scale_x_reverse()+

  geom_path(data = LIBS %>% filter(specimen=="C18",sample_mode=="trace"),
            aes(x = x/ 1000, y = y/ 1000),
            col = "red",
            linetype="dashed",
            size=0.7)+
    ylab("y-axis [mm]")+ggtitle("C18")+
  xlab("x-axis [mm]")+labs(col = "Mg/Ca [a.u.]")



C18_Line <- LIBS %>% 
  filter(specimen=="C18",
         sample_mode == "trace") %>%
  ggplot() +
  aes(dist,
      mg_ca,
      col = mg_ca) +
  geom_path(show.legend=FALSE,size=1) +
  scale_color_viridis(option = "G",
                      begin = 0.5,
                      end = 0.95) +
  ylab("Mg/Ca [a.u.]") +
  xlab("Distance to edge [mm]")

  

```

## Final Plot Fig. 6
```{r}

C16_Map+C17_Map+C18_Map+C16_Line+C17_Line+C18_Line+plot_annotation(
                          title="North",
                          tag_levels = "a"
                          )+plot_layout(ncol = 3,widths = c(0.5,0.5),guides = "collect")

# ggsave(here("Figures", "Figure 6.png"), units = "cm", width = 29.7, height = 14.85)


```



# Figure 7 - Sequential stable isotope data
```{r}
Isotopes %>% 
  filter(element=="O") %>% 
ggplot()+
  aes(
    x=sample_seq,
    y=VPDB,
  )+
  geom_path()+
  geom_point() +
  ylab(expression("δ" ^ "18" * "O‰ [VPDB]"))+
  # theme(legend.title = element_text(colour="white"))+
  # scale_linetype_discrete(
    # limits = c("O", "C"),
    # labels = c(expression("δ" ^ "18" * "O"), expression("δ" ^ "13" * "C")))+
  # theme(legend.position = c(0.4,0.5),
  #       legend.key.size = unit(1, "cm"))+
  facet_wrap(~specimen) +
  theme(strip.text = element_text(size=15, face="bold", hjust=0),
        strip.background = element_rect(colour="white", fill="white")) +
  xlab("Sample sequence")


# ggsave(here("Figures", "Figure 7.png"), units = "cm", width = 29.7, height = 14.85)


```

# Figure 8 - Correlation graphs of the two sequentially sampled shells for SI analysis
## Turn SST data into δ^18^O values
```{r}
SST_O <- 
  SST %>% select(date,sst) %>%
  mutate(date=as.numeric(date)-16071) %>% #needs numeric value for matrix
  as.matrix() %>% # next function needs matrix
  d18O_model(d18Ow=0.55,mineral="calcite") %>% # δ18Owater is 0.5 based on δ18O=( 0.25- S)−8.2 and salinity (S) is around 35 psu.
  as_tibble() %>% 
  rename(date=V1,d18O=V2) %>% 
  mutate(date=as.Date.numeric(date,origin = "2014-01-02")) #put dates back
```

## C16
#### Warp data
```{r}

SeqC16_O <- Isotopes %>% 
  filter(specimen=="C16",
         element=="O") %>% 
  rename(O=VPDB) %>% 
  na.omit() %>% 
  arrange(desc(row_number())) %>% #flip the data so the past is facing the other way 
  pull(O)
  
SeqSST_O<- SST_O %>% 
  filter(date>"2014-02-25",
         date<"2017-06-02") %>% 
  pull(d18O)


alignment_C16 <- dtw(SeqC16_O,SeqSST_O,
                 keep=TRUE,
                 step.pattern=symmetric2)
    
# Uncomment to see alignment
# plot(alignment_C16,
#      type = "twoway",
#      offset = 3)
```

#### Adjust data

```{r}

Al_C16_O <- Isotopes %>%
  filter(specimen=="C16",
         element=="O") %>%
  na.omit() %>%
  mutate(x=43:1,
         proxy="d18Os") %>%
  rename(value=VPDB) %>%
  select(x,value,proxy)


Al_C16_SST_O <- SST_O %>%
  filter(date>"2014-02-25",
         date<"2017-06-02") %>% 
  pull(d18O) %>%
  .[alignment_C16$index2] %>% 
  as.data.frame() %>% 
  clean_names() %>% 
  rename(value=x) %>% 
  cbind(x=alignment_C16$index1) %>% 
    mutate(proxy="d18Oest")


C16_SST_O_aligned <- rbind(Al_C16_O,Al_C16_SST_O) %>% 
  group_by(x,proxy) %>%
  summarise(value=mean(value)) %>% #this unifies points that were given the same location
  ungroup() %>%
  mutate(value=round(value+0.000001*row_number(),5)) %>% #this adds some error to data points that are similar but far away from each other and prevents them to be grouped as one in the grouping step below
  group_by(proxy,value) %>% # this unifies extra points that were added to fill gaps in the record
  summarise(x=mean(x)) 


Al_C16_O_plot <- C16_SST_O_aligned %>%
  filter(proxy=="d18Os") %>% 
ggplot()+
  aes(x=x,
      y=value)+
  geom_line()+
  xlab("arbitrary distance")+
  ylab("‰ VPDB")+
  ggtitle("Shell data")

Al_C16_SST_O_plot <- C16_SST_O_aligned %>%
  filter(proxy=="d18Oest") %>% 
ggplot()+
  aes(x=x,
      y=value)+
  geom_line()+
  xlab("arbitrary distance")+
    ylab("‰ VPDB")+
    ggtitle("Water data")



# Uncomment to see aligned data
# Al_C16_O_plot/Al_C16_SST_O_plot
``` 




### C17
#### Warp data

```{r}

SeqC17_O <- Isotopes %>% 
  filter(specimen=="C17",
         element=="O") %>% 
  rename(O=VPDB) %>% 
  na.omit() %>% 
  arrange(desc(row_number())) %>% #flip the data so the past is facing the other way 
  pull(O)
  

alignment_C17 <- dtw(SeqC17_O,SeqSST_O,
                 keep=TRUE,
                 step.pattern=symmetric2)
# Uncomment to see alignment           
# plot(alignment_C17,
#      type = "twoway",
#      offset = 3)
```

#### Adjust data

```{r}

Al_C17_O <- Isotopes %>%
  filter(specimen=="C17",
         element=="O") %>%
  na.omit() %>%
  mutate(x=46:1,
         proxy="d18Os") %>%
  rename(value=VPDB) %>%
  select(x,value,proxy)

  
Al_C17_SST_O <- SST_O %>%
  filter(date>"2014-02-25",
         date<"2017-06-02") %>% 
  pull(d18O) %>%
  .[alignment_C17$index2] %>% 
  as.data.frame() %>% 
  clean_names() %>% 
  rename(value=x) %>% 
  cbind(x=alignment_C17$index1) %>% 
    mutate(proxy="d18Oest")


C17_SST_O_aligned <- rbind(Al_C17_O,Al_C17_SST_O) %>% 
  group_by(x,proxy) %>%
  summarise(value=mean(value)) %>% #this unifies points that were given the same location
  ungroup() %>%
  mutate(value=round(value+0.000001*row_number(),5)) %>% #this adds some error to data points that are similar but far away from each other and prevents them to be grouped as one in the grouping step below
  group_by(proxy,value) %>% # this unifies extra points that were added to fill gaps in the record
  summarise(x=mean(x)) 


Al_C17_O_plot <- C17_SST_O_aligned %>%
  filter(proxy=="d18Os") %>% 
ggplot()+
  aes(x=x,
      y=value)+
  geom_line()+
  xlab("arbitrary distance")+
  ylab("‰ VPDB")+
  ggtitle("Shell data")

Al_C17_SST_O_plot <- C17_SST_O_aligned %>%
  filter(proxy=="d18Oest") %>% 
ggplot()+
  aes(x=x,
      y=value)+
  geom_line()+
  xlab("arbitrary distance")+
    ylab("‰ VPDB")+
    ggtitle("Water data")



# Uncomment to see alignment
# Al_C17_O_plot/Al_C17_SST_O_plot
``` 

## SST to d18O correlations
### C16
```{r}
# C16
df_C16_SST_O <- C16_SST_O_aligned %>%
  pivot_wider(names_from = "proxy",
              values_from = "value") %>% 
  na.omit() %>% 
  select(d18Os,d18Oest)

LM_C16_SST_O <- lm(d18Os ~ d18Oest,
         data=df_C16_SST_O,
         )
R2_C16_SST_O <- round(summary(LM_C16_SST_O)[[8]],2)

P_C16_SST_O <- ifelse(lmp(LM_C16_SST_O)<0.001,"p<0.001",
                ifelse(lmp(LM_C16_SST_O)<0.01,"p<0.01",
                       ifelse(lmp(LM_C16_SST_O)<0.05,"p<0.05","ns!!!")))

Stat_C16_SST_O <- paste0("R²=", R2_C16_SST_O,", ",P_C16_SST_O)



Fig8_a <- C16_SST_O_aligned %>%
  pivot_wider(names_from = "proxy",
              values_from = "value") %>%
  ggplot() +
  aes(x = d18Os,
      y = d18Oest) +
  geom_point() +
  # xlim(-0.5, 2.5) +
  # ylim(0.2, 0.5) +
  geom_smooth(method = "lm",
              se = FALSE,
              col="black",
              linetype="dashed",
              size=0.5) +
ylab(expression("δ"^"18"*"O"[Water]*" [VPDB]")) +
  xlab(expression("δ" ^ "18" * "O"[Shell]*" [VPDB]"))+
  ggtitle("C16",#"Correlation graph for C17 d18O and Mg/Ca",
          subtitle = Stat_C16_SST_O)



```

### C17
```{r}
# C17
df_C17_SST_O <- C17_SST_O_aligned %>%
  pivot_wider(names_from = "proxy",
              values_from = "value") %>% 
  na.omit() %>% 
  select(d18Os,d18Oest)

LM_C17_SST_O <- lm(d18Os ~ d18Oest,
         data=df_C17_SST_O,
         )
R2_C17_SST_O <- round(summary(LM_C17_SST_O)[[8]],2)

P_C17_SST_O <- ifelse(lmp(LM_C17_SST_O)<0.001,"p<0.001",
                ifelse(lmp(LM_C17_SST_O)<0.01,"p<0.01",
                       ifelse(lmp(LM_C17_SST_O)<0.05,"p<0.05","ns!!!")))

Stat_C17_SST_O <- paste0("R²=", R2_C17_SST_O,", ",P_C17_SST_O)


Fig8_b <- C17_SST_O_aligned %>%
  pivot_wider(names_from = "proxy",
              values_from = "value") %>%
  ggplot() +
  aes(x = d18Os,
      y = d18Oest) +
  geom_point() +
  # xlim(-0.5, 2.5) +
  # ylim(0.2, 0.5) +
  geom_smooth(method = "lm",
              se = FALSE,
              col="black",
              linetype="dashed",
              size=0.5) +
  ylab(expression("δ" ^ "18" * "O"[Water]*" [VPDB]"))+
  xlab(expression("δ" ^ "18" * "O"[Shell]*" [VPDB]"))+
  ggtitle("C17",#"Correlation graph for C17 d18O and Mg/Ca",
          subtitle = Stat_C17_SST_O)



```
## Final Plot Figure 8
```{r}
Fig8_a+Fig8_b
# ggsave(here("Figures", "Figure 8.png"), units = "cm", width = 29.7, height = 14.85)

```

# Figure 9 - Comparison of climatic and geochemical datasets of C16 and C17
## SST
```{r}

season_colors <- c(
  Summer = "#FDBE85", # Warm, light orange for Summer
  Autumn = "#F8766D", # Rich, earthy orange for Autumn
  Winter = "#6BAED6", # Cool, soft blue for Winter
  Spring = "#31A354"  # Fresh, vibrant green for Spring
)

SST_plot <-
  SST %>% 
  ggplot()+
  aes(x=date,y=sst,col=season,group=year) +
  geom_line()+
  ylab("SST ºC")+
    scale_color_manual(values = season_colors)+
theme(legend.position = "bottom", legend.justification = "center")+
  theme(legend.margin = margin(t = -45, unit = "pt"))+
  labs(y="SST ºC",x="",col="")+
  geom_vline(xintercept = as.numeric(as.Date("2014-02-25")), linetype=2)+
  geom_vline(xintercept = as.numeric(as.Date("2015-02-25")), linetype=2)+
  geom_vline(xintercept = as.numeric(as.Date("2016-02-10")), linetype=2)+
  geom_vline(xintercept = as.numeric(as.Date("2017-03-20")), linetype=2)

```


## C16
```{r}
C16_O_plot <- Isotopes %>% 
  filter(element=="O",
         specimen=="C16") %>%
  na.omit() %>% 
  ggplot()+
  aes(x=sample_seq,y=VPDB)+
  geom_line()+
  geom_vline(xintercept = 43, linetype=2)+
  geom_vline(xintercept = 34, linetype=2)+
  geom_vline(xintercept = 28, linetype=2)+
  geom_vline(xintercept = 4, linetype=2)+
  ylim(2.5,-0.5)+
  ylab(expression("δ" ^ "18" * "O [VPDB]"))+
  xlab("sample number")+
  scale_x_reverse()

C16_Mg_Ca_plot <-
  LIBS %>% 
   filter(specimen=="C16",
         sample_mode == "trace") %>%
  ggplot() +
  aes(x = dist,
      y = mg_ca) +
  geom_path() +
  scale_x_reverse()+
  xlab("Distance to Edge [mm]")+
  ylab("Mg/Ca ratio")+
  geom_vline(xintercept = 0.3, linetype=2)+
  geom_vline(xintercept = 3.5, linetype=2)+
  geom_vline(xintercept = 8.9, linetype=2)+
  geom_vline(xintercept = 6, linetype=2)+
  geom_vline(xintercept = 4.4, linetype=2)+
  annotate("rect", xmin = 0, xmax = 6.05, ymin = 0.25, ymax = 0.47,
  alpha = .2)+
  annotate("text",x=1.1,y=0.49,label=as.character(expression("extent of δ" ^ "18" * "O record")), alpha=.6,parse=TRUE,size=2.5)


Fig9_a <- SST_plot/C16_O_plot/C16_Mg_Ca_plot+plot_annotation(title = "C16")

```

## C17
```{r}

C17_O_plot <- Isotopes %>% 
  filter(element=="O",
         specimen=="C17") %>% 
  ggplot()+
  aes(x=sample_seq,y=VPDB)+
  geom_line()+
  geom_vline(xintercept = 10, linetype=2)+
  geom_vline(xintercept = 37, linetype=2)+
  geom_vline(xintercept = 44, linetype=2)+
  geom_vline(xintercept = 51, linetype=2)+
  ylim(2.5,-0.5)+
  ylab(expression("δ" ^ "18" * "O [VPDB]"))+
  xlab("sample number")+
  scale_x_reverse()

C17_Mg_Ca_plot <-
  LIBS %>% 
   filter(specimen=="C17",
         sample_mode == "trace") %>% 
  ggplot() +
  aes(x = dist,
      y = mg_ca) +
  geom_path() +
  scale_x_reverse()+
  xlab("Distance to Edge [mm]")+
  ylab("Mg/Ca ratio")+
  geom_vline(xintercept = 0.8, linetype=2)+
  geom_vline(xintercept = 3.5, linetype=2)+
  geom_vline(xintercept = 8.6, linetype=2)+
  geom_vline(xintercept = 5.4, linetype=2)+
  geom_vline(xintercept = 4.7, linetype=2)+
  annotate("rect", xmin = 0, xmax = 5.6, ymin = 0.25, ymax = 0.61,
  alpha = .2)+
  annotate("text",x=1.65,y=0.64,label=as.character(expression("extent of δ" ^ "18" * "O record")), alpha=.6,parse=TRUE,size=2.5)



Fig9_b <- SST_plot/C17_O_plot/C17_Mg_Ca_plot+plot_annotation(title = "C17")

```

## Final Plot for Figure 9
```{r}
Fig9_a|Fig9_b

# ggsave(here("Figures", "Figure 9.png"), units = "cm", width = 29.7, height = 14.85)
```

# Figure 10 - Correlation graphs of the two sequentially sampled shells for SI analysis with Mg/Ca data
## C16
### Warp data

```{r fig.height=4}

SeqC16_O <- Isotopes %>% 
  filter(specimen=="C16",
         element=="O") %>% 
  mutate(O=scale(VPDB)) %>% 
  na.omit() %>% 
  pull(O)
  
SeqC16_mg_ca <- LIBS %>% 
  filter(specimen=="C16",
         sample_mode=="trace",
         dist<5.5) %>% # this value was chosen as an estimate to shorten the LIBS data to the length of the isotope data
    mutate(mg_ca=scale(mg_ca)) %>% 
  pull(mg_ca) 

alignment_C16 <- dtw(SeqC16_O*-1,SeqC16_mg_ca,
                 keep=TRUE,
                 step.pattern=symmetric2)
           
# Uncomment to see alignment of datasets
# plot(alignment_C16,
#      type = "twoway",
#      offset = 3)
```

### Adjust data
```{r}

Al_C16_O <- Isotopes %>% 
  filter(specimen=="C16",
         element=="O") %>%  
  na.omit() %>% 
  mutate(x=1:43,
         proxy="O") %>% 
  rename(value=VPDB) %>% 
  select(x,value,proxy)
  
Al_C16_mg_ca <- LIBS %>% 
  filter(sample_mode=="trace",
         specimen=="C16",
         dist<5.5) %>%
  pull(mg_ca) %>%
  .[alignment_C16$index2] %>% 
  as.data.frame() %>% 
  clean_names() %>% 
  rename(value=x) %>% 
  cbind(x=alignment_C16$index1) %>% 
    mutate(proxy="mg_ca")


C16_aligned <- bind_rows(Al_C16_O,Al_C16_mg_ca) %>% 
  group_by(x,proxy) %>%
  summarise(value=mean(value)) %>% #this unifies points that were given the same location
  ungroup() %>%
  mutate(value=round(value+0.000001*row_number(),5)) %>% #this adds some error to data points that are similar but far away from each other and prevents them to be grouped as one in the grouping step below
  group_by(proxy,value) %>% # this unifies extra points that were added to fill gaps in the record
  summarise(x=mean(x))

# Uncomment to see alignment of both adjusted datasets
# Al_C16_O_map <- C16_aligned %>%
#   filter(proxy=="O") %>% 
# ggplot()+
#   aes(x=x,
#       y=value)+
#   geom_line()+
#   scale_y_reverse()+
#   xlab("arbitrary distance")+
#   ylab("inverse VPDB")
# 
# Al_C16_mg_ca_map <- C16_aligned %>%
#   filter(proxy=="mg_ca") %>% 
# ggplot()+
#   aes(x=x,
#       y=value)+
#   geom_line()+
#   xlab("arbitrary distance")+
#     ylab("Mg/Ca ratio")
#   # ylim(0.,0.4)
# 
# 
# 
# Al_C16_O_map/Al_C16_mg_ca_map


```

### Correlation of d18O to Mg/Ca
```{r}
# C16
df_C16 <- C16_aligned %>%
  pivot_wider(names_from = "proxy",
              values_from = "value") %>% 
  select(O,mg_ca)

LM_C16 <- lm(O ~ mg_ca,
         data=df_C16,
         )
R2_C16 <- round(summary(LM_C16)[[8]],2)

P_C16 <- ifelse(lmp(LM_C16)<0.001,"p<0.001",
                ifelse(lmp(LM_C17)<0.01,"p<0.01",
                       ifelse(lmp(LM_C17)<0.05,"p<0.05","ns!!!")))

Stat_C16 <- paste0("R²=", R2_C16,", ",P_C16)

Fig10_a <- C16_aligned %>%
  pivot_wider(names_from = "proxy",
              values_from = "value") %>%
  ggplot() +
  aes(x = O,
      y = mg_ca) +
  geom_point() +
  xlim(-0.5, 2.5) +
  ylim(0.2, 0.5) +
  geom_smooth(method = "lm",
              se = FALSE,
              col="black",
              linetype="dashed",
              size=0.5) +
  ylab("Mg/Ca Ratio") +
  xlab(expression("δ" ^ "18" * "O [VPDB]"))+
  ggtitle("C16",#"Correlation graph for C16 d18O and Mg/Ca",
          subtitle = Stat_C16)
```


## C17
### Warp data

```{r fig.height=4}

SeqC17_O <- Isotopes %>% 
  filter(specimen=="C17",
         element=="O") %>% 
  mutate(O=scale(VPDB)) %>% 
  na.omit() %>% 
  pull(O)
  
SeqC17_mg_ca <- LIBS %>% 
  filter(specimen=="C17",
         sample_mode=="trace",
         dist<5.2) %>% # this value was chosen as an estimate to shorten the LIBS data to the length of the isotope data
    mutate(mg_ca=scale(mg_ca)) %>% 
  pull(mg_ca) 

alignment_C17 <- dtw(SeqC17_O*-1,SeqC17_mg_ca,
                 keep=TRUE,
                 step.pattern=symmetric2)
           
# Uncomment to see alignment of datasets
plot(alignment_C17,
     type = "twoway",
     offset = 3)
```

### Adjust data
```{r}

Al_C17_O <- Isotopes %>% 
  filter(specimen=="C17",
         element=="O") %>%  
  na.omit() %>% 
  mutate(x=1:46,
         proxy="O") %>% 
  rename(value=VPDB) %>% 
  select(x,value,proxy)
  
Al_C17_mg_ca <- LIBS %>% 
  filter(sample_mode=="trace",
         specimen=="C17",
         dist<5.2) %>%
  pull(mg_ca) %>%
  .[alignment_C17$index2] %>% 
  as.data.frame() %>% 
  clean_names() %>% 
  rename(value=x) %>% 
  cbind(x=alignment_C17$index1) %>% 
    mutate(proxy="mg_ca")


C17_aligned <- bind_rows(Al_C17_O,Al_C17_mg_ca) %>% 
  group_by(x,proxy) %>%
  summarise(value=mean(value)) %>% #this unifies points that were given the same location
  ungroup() %>%
  mutate(value=round(value+0.000001*row_number(),5)) %>% #this adds some error to data points that are similar but far away from each other and prevents them to be grouped as one in the grouping step below
  group_by(proxy,value) %>% # this unifies extra points that were added to fill gaps in the record
  summarise(x=mean(x))

# Uncomment to see alignment of both adjusted datasets
# Al_C17_O_map <- C17_aligned %>%
#   filter(proxy=="O") %>%
# ggplot()+
#   aes(x=x,
#       y=value)+
#   geom_line()+
#   scale_y_reverse()+
#   xlab("arbitrary distance")+
#   ylab("inverse VPDB")
# 
# Al_C17_mg_ca_map <- C17_aligned %>%
#   filter(proxy=="mg_ca") %>%
# ggplot()+
#   aes(x=x,
#       y=value)+
#   geom_line()+
#   xlab("arbitrary distance")+
#     ylab("Mg/Ca ratio")
#   # ylim(0.,0.4)
# 
# 
# 
# Al_C17_O_map/Al_C17_mg_ca_map


```

### Correlation of d18O to Mg/Ca
```{r}
# C17
df_C17 <- C17_aligned %>%
  pivot_wider(names_from = "proxy",
              values_from = "value") %>% 
  select(O,mg_ca)

LM_C17 <- lm(O ~ mg_ca,
         data=df_C17,
         )
R2_C17 <- round(summary(LM_C17)[[8]],2)

P_C17 <- ifelse(lmp(LM_C17)<0.001,"p<0.001",
                ifelse(lmp(LM_C17)<0.01,"p<0.01",
                       ifelse(lmp(LM_C17)<0.05,"p<0.05","ns!!!")))

Stat_C17 <- paste0("R²=", R2_C17,", ",P_C17)

Fig10_b <- C17_aligned %>%
  pivot_wider(names_from = "proxy",
              values_from = "value") %>%
  ggplot() +
  aes(x = O,
      y = mg_ca) +
  geom_point() +
  xlim(-0.5, 2.5) +
  ylim(0.2, 0.5) +
  geom_smooth(method = "lm",
              se = FALSE,
              col="black",
              linetype="dashed",
              size=0.5) +
  ylab("Mg/Ca Ratio") +
  xlab(expression("δ" ^ "18" * "O [VPDB]"))+
  ggtitle("C17",#"Correlation graph for C17 d18O and Mg/Ca",
          subtitle = Stat_C17)
```

## Final Plot for Figure 10
```{r}
Fig10_a+Fig10_b

# ggsave(here("Figures", "Figure 10.png"), units = "cm", width = 29.7, height = 14.85)

```

